Proposed Solution

### 2017-04-17 OOPSLA17/secs/trans.tex

\para{Probabilistic Types}
%\label{sec:ptypes}
An additional challenge is that it is not always possible to fully determine the type of key based on a single example value. 
For this reason, we introduce \textit{probabilistic types}, as contrasted with \textit{basic types}.
In \app, the set of basic types contains strings, file paths, integers, sizes, and Booleans. 
Taking the configuration \ref{lst:file1.cnf}, we can assume {\tt foo} is a Boolean type by the grammar of MySQL,
  but the keyword {\tt bar} could be many types.
If we choose the type based on the first example, {\tt bar} will be a integer type.
If we choose a type that fits all examples, {\tt bar} will be a string.
However the correct classification needed is a Boolean type.
For illustrative purposes, let us assume that there is a critical rule we must learn where tow Boolean keywords should always have the same values, e.g. $eq(\texttt{foo},\texttt{bar})$.
If we take {\tt bar}$::int$, we do not learn the above rule, nor do we learn this rule with {\tt bar}$::string$ - only with {\tt bar}$::bool$ is the rule is valid.
To resolve this ambiguity, and choose the best type, the translator assigns a distribution of types to a keyword based on examples from the training set of configuration files (denoted $\trainingSet=\{C\}$).

@@

{
\setlength{\abovecaptionskip}{-.05pt}
\setlength{\belowcaptionskip}{-15pt}
\begin{figure}
\begin{mathpar}
%\hspace{-1cm} %no idea why, but this is better than \centering here
\inferrule* [Right=ptype]
{c_{int} = |\{\forall C \in \trainingSet.\ \forall (k,v) \in C.\ v \in \mathbb{Z}\}| \\\\ c_{bool} = |\{\forall C \in \trainingSet.\ \forall (k,v) \in C.\ v \in \{0,1,ON,OFF\}\}| }
{k : \ptype[int = c_{int}, bool = c_{bool}]}
\and\\
\inferrule* [Right=int]
{k:\ptype \\ p_{int}(\ptype)}
{k :: int}
\and
\inferrule* [Right=bool]
{k:\ptype \\ p_{bool}(\ptype)}
{k :: bool}
\and\\
\inferrule* [Right=eq\_rule]
{k_1 :: \tau \\ k_2 :: \tau}
{eq(k_1,k_2) :: Rule}
\and
\inferrule* [Right=order\_rule]
{k1,k2 \in C \\ k1 \neq k2}
{ord(k1,k2) :: Rule}
\end{mathpar}
\caption{Type judgments for a probabilistic type system with $\mathcal{T} = \{bool,int\}$ and an equality and ordering rule}
\label{fig:ptypes}
\end{figure}
}

In order to define resolution predicates, we use the notation $|\tau_i\ptype|$ to select $c_i$ from a $\ptype \in \tilde{\mathcal{T}}$.
As an example, for the sample training set provided in Figure~\ref{fig:tset}, we might choose to set $p_{bool}(\ptype)= |bool_{\ptype}| \geq 3 \land |int_{\ptype}| \leq 1$ and $p_{int}(\ptype)= |int\ptype| \geq 3$.
We then can run inference on that sample training set to derive a probabilistic type \texttt{bar}$:[bool=3,int=1]$.
This is then resolved, as required in our earlier example, to \texttt{bar}$::bool$.

Note that a user may pick predicates for probabilistic type resolution that result in overlapping inference rules.
For example, if a user instead picked $p_{int}(\ptype)= |int\ptype| \geq 1$, {\scriptsize INT} would overlap with {\scriptsize BOOL}.
This means that types are not unique in this system, i.e. $k:\ptype \land k:\ptype' \centernot\implies \ptype = \ptype'$.

>>>>

\para{Probabilistic Types}
%\label{sec:ptypes}
An additional challenge is that it is not always possible to fully determine the type of key based on a single example value. 
For this reason, we introduce \textit{probabilistic types}, as contrasted with \textit{basic types}.
In \app, the set of basic types contains strings, file paths, integers, sizes, and Booleans. 
Taking the configuration \ref{lst:file1.cnf}, we can assume {\tt foo} is a Boolean type by the grammar of MySQL,
  but the keyword {\tt bar} could be many types.
If we choose the type based on the first example, {\tt bar} will be a integer type.
If we choose a type that fits all examples, {\tt bar} will be a string.
However the correct classification needed is a Boolean type.
For illustrative purposes, let us assume that there is a critical rule we must learn where tow Boolean keywords should always have the same values, \eg $eq(\texttt{foo},\texttt{bar})$.
If we take {\tt bar}$::int$, we do not learn the above rule, nor do we learn this rule with {\tt bar}$::string$ - only with {\tt bar}$::bool$ is the rule is valid.
To resolve this ambiguity, and choose the best type, the translator assigns a distribution of types to a keyword based on examples from the training set of configuration files (denoted $\trainingSet=\{C\}$).

@@

{
\setlength{\abovecaptionskip}{-.05pt}
\setlength{\belowcaptionskip}{-15pt}
\begin{figure}
\begin{mathpar}
%\hspace{-1cm} %no idea why, but this is better than \centering here
\inferrule* [Right=ptype]
{c_{int} = |\{\forall C \in \trainingSet.\ \forall (k,v) \in C.\ v \in \mathbb{Z}\}| \\\\ c_{bool} = |\{\forall C \in \trainingSet.\ \forall (k,v) \in C.\ v \in \{0,1,ON,OFF\}\}| }
{k : \ptype[int = c_{int}, bool = c_{bool}]}
\and\\
\inferrule* [Right=int]
{k:\ptype \\ p_{int}(\ptype)}
{k :: int}
\and
\inferrule* [Right=bool]
{k:\ptype \\ p_{bool}(\ptype)}
{k :: bool}
\and\\
\inferrule* [Right=eq\_rule,before=\hspace{-1.5cm}]
{k_1 :: \tau \\ k_2 :: \tau}
{eq(k_1,k_2) :: Rule}
\and
\inferrule* [Right=order\_rule,before=\hspace{-1cm}]
{k1,k2 \in C \\ k1 \neq k2}
{ord(k1,k2) :: Rule}
\end{mathpar}
\caption{Type judgments for a probabilistic type system with $\mathcal{T} = \{bool,int\}$ and an equality and ordering rule}
\label{fig:ptypes}
\end{figure}
}

In order to define resolution predicates, we use the notation $|\tau_{i\ptype}|$ to select $c_i$ from a $\ptype \in \tilde{\mathcal{T}}$.
As an example, for the sample training set provided in Figure~\ref{fig:tset}, we might choose to set $p_{bool}(\ptype)= |bool_{\ptype}| \geq 3 \land |int_{\ptype}| \leq 1$ and $p_{int}(\ptype)= |int_{\ptype}| \geq 3$.
We then can run inference on that sample training set to derive a probabilistic type \texttt{bar}$:[bool=3,int=1]$.
This is then resolved, as required in our earlier example, to \texttt{bar}$::bool$.

Note that a user may pick predicates for probabilistic type resolution that result in overlapping inference rules.
For example, if a user instead picked $p_{int}(\ptype)= |int_{\ptype}| \geq 1$, this predicate would overlap with the $p_{bool}$ predicate.
This means that basic types are not unique in this system, \ie $k::\tau \land k::\tau' \centernot\implies \tau = \tau'$.
